### 2020-10-09 report/report.tex
Scaffold LaTeX report

\documentclass[]{article}

%opening
\title{Detecting Causal Structure in Time Series Data}
\author{Lawrence Tray \\ Ioannis Kontoyiannis}

%packages
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

%package setup
\graphicspath{{./img/}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%custom commands
\newcommand{\dft}{\mathcal{F}}
\newcommand{\idft}{\mathcal{F}^{-1}}
\newcommand{\Xcal}{\mathcal{X}}
\newcommand{\cmplx}{\mathbb{C}}
\newcommand{\figwidth}{0.55\linewidth}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}

\end{document}