### 2020-11-11 reading/reading.tex
Scaffold bibliography

\documentclass[]{article}

%opening
\title{Jormungandr - Reading Notes}
\author{Lawrence Tray \\ Ioannis Kontoyiannis}

%packages
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools}

\usepackage[backend=bibtex]{biblatex}
\addbibresource{sources.bib}

\begin{document}

\maketitle

\section{Initial Reading}

I read \cite{Abbe}

\printbibliography

\end{document}