### 2021-05-24 cued/checklist.tex

\section*{Checklist}

\begin{enumerate}

	\item For all authors...
	\begin{enumerate}
		\item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
		\answerYes{See experiments section \ref{sec:experiments}}
		\item Did you describe the limitations of your work?
		\answerYes{See conclusion \ref{sec:conclusion} and experiments \ref{sec:experiments} sections}
		\item Did you discuss any potential negative societal impacts of your work?
		\answerYes{See conclusion \ref{sec:conclusion}}
		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
		\answerYes{}
	\end{enumerate}

	\item If you are including theoretical results...
	\begin{enumerate}
		\item Did you state the full set of assumptions of all theoretical results?
		\answerYes{See inference section \ref{sec:inference}}
		\item Did you include complete proofs of all theoretical results?
		\answerYes{See inference section \ref{sec:inference}}
	\end{enumerate}

	\item If you ran experiments...
	\begin{enumerate}
		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
		\answerYes{See supplementary code}
		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
		\answerYes{See appendix \ref{appdx:hyperparams}}
		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
		\answerYes{See table \ref{tab:results}}
		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
		\answerYes{See appendix \ref{appdx:imp-details}}
	\end{enumerate}

	\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
	\begin{enumerate}
		\item If your work uses existing assets, did you cite the creators?
		\answerYes{See section experiments \ref{sec:experiments}}
		\item Did you mention the license of the assets?
		\answerYes{See section inference \ref{sec:inference}}
		\item Did you include any new assets either in the supplemental material or as a URL?
		\answerYes{Supplementary material}
		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
		\answerYes{Referred to original papers}
		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
		\answerYes{Referred to original papers}
	\end{enumerate}

	\item If you used crowdsourcing or conducted research with human subjects...
	\begin{enumerate}
		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
		\answerNA{}
		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
		\answerNA{}
		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
		\answerNA{}
	\end{enumerate}

\end{enumerate}