### 2021-05-24 cued/prelim.tex

\section{Preliminaries}

We first need a model for community-like structure in a network. For this we adopt the stochastic block model (SBM) - widely used across academia. Each node in the graph belongs to a unique community called a block. The probability that two nodes are connected depends only on the block memberships of each. Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}. To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation (DC-SBM).

\begin{definition}[Microcanonical DC-SBM]
	\label{defn:microcan-dc-sbm}
	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph. The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{1, 2 \dots K\}$ to compactly define a set of $K$ indices. Clearly, $[K]$ is only defined for $K \in \Integers^+$.}
	where $B \in \Integers^{+}$ is the number of non-empty blocks.
	Let $e \in (\Integers_0^+)^{N \times N}$ be the matrix of edge counts between blocks. $e_{rs}$ is then the number of edges from block $r$ onto block $s$ -- or twice that number if $r=s$. For undirected graphs, $e$ is symmetric.
	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph. $k_i$ is then the degree of vertex $i$. The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
	%
	\begin{equation}
		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
	\end{equation}
	%
	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence the microcanonical moniker. Specifically, $A$ must satisfy the following:
	%
	\begin{equation}
		e_{rs} = \sum_{i, j \in [N]} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \quad \forall r,s \in [B]
		\quad \textrm{and} \quad
		k_i = \sum_{j \in [N]} A_{ij} \quad \forall i \in [N]
		\label{eqn:sbm-constraints}
	\end{equation}
\end{definition}