### 2021-05-24 cued/covid.tex
Font-size and title page

\section{Covid-19 disruption}

Please provide details of disrupted project work, including dates when the work was planned to take place and
reasons for the disruption (e.g. inaccessible equipment, illness).
If your work was disrupted, to what extent was it possible to redefine elements of your project so as to achieve
alternative goals? Please comment on the success, or otherwise, of the Covid-19 contingency plan that you
formulated last summer and documented on your Project Planning Form.
This appendix should not exceed one page in length.