### 2021-05-24 cued/covid.tex
Covid and risk retro

\section{Covid-19 disruption}

Thankfully this project was not directly disrupted by Covid-19 as it was always going to be heavily theoretical and computational. The only disruption to speak of was the shift of supervisor meetings to online. Though shared whiteboards were alien at first, we grew to learn how to use them to effectively communicate ideas.

I was ill with Covid-19 for 2 weeks in October but this did not majorly impact my ability to work. The Covid-19 contingency plan formulated last summer and documented on the Project Planning Form just involved switching supervisions to an online format.