### 2021-05-24 cued/risk.tex

\section{Risk assessment retrospective}

As experimental work was entirely computer-based the risks encountered were not what one might typically associate with a more hands-on project: heavy machinery, dangerous chemicals etc. Most of the risks identified would be as a result of an improper working environment. These could be properly managed by ensuring the working environment felt comfortable. An external monitor went some way to alleviating the eye-strain of coding on a small screen for extended periods. Furthermore, an external mouse is far kinder on the wrist than a track-pad on laptops. There were no unexpected hazards encountered that were not identified at the start of the project.