### 2021-05-27 nips/introduction.tex

\section{Introduction}

An important characteristic of many real-world networks is that they exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
Finding ways to recover the latent communities from the observed graphs is an important
task in many applications, includng
graph/network compression \cite{cluster-compression} and link prediction in 
incomplete networks \cite{link-prediction}.

In this work, we
restrict our attention to vertex-labelled networks, and we
refer to the vertex labels as {\em features}. A common goal is to determine whether a given feature impacts graphical structure. To answer this from a Bayesian perspective requires the use of a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}. This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each. Numerous variants of this 
model have been considered -- the most popular ones being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}. Effectively, these extend the model to allow each vertex to belong to multiple blocks simultaneously. However, one drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process. Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.

To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks. Nevertheless, this approach is limited in that it can only consider one feature at a time. This makes it difficult to rank-order the features by  magnitude of impact. Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit. We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.

Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM). This is an extension of the SBM to labelled networks. We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator, and to describe how the 
sampled FFBM parameters can be interpreted in order to determine which features have the largest impact on overall graphical structure.