### 2021-05-31 cued/tech-abstract.tex
Modifications

\section*{Technical Abstract}
Labelled networks form a very common and important class of data,
naturally appearing in numerous applications in science and engineering.
A typical inference goal is to determine how the vertex labels
(or {\em features}) affect the network's graph structure.

A standard approach has been to partition the network into blocks grouped
by distinct values of the feature of interest. A block-based random
graph model -- typically a variant of the stochastic block model --
is then used to test for evidence of asymmetric behaviour within these
feature-based communities. Nevertheless, the resulting communities
often do not produce a natural partition of the graph.

In this work, we introduce a new generative model, the feature-first block model (FFBM),
which is more effective at describing vertex-labelled undirected
graphs and also facilitates the use of richer queries on labelled networks.
We develop a Bayesian framework for inference with this model,
and we present a method to efficiently sample from the posterior
distribution of the FFBM parameters. The FFBM's structure is kept
deliberately simple to retain easy interpretability of the parameter
values.

We apply the proposed methods to a variety of network data
to extract the most important features along which the vertices
are partitioned. The main advantages of the proposed approach are
that the whole feature-space is used automatically and that features
can be rank-ordered implicitly according to impact. Any features
that do not significantly impact the high-level structure can be
discarded to reduce the problem dimension. In cases where the vertex
features available do not readily explain the community structure
in the resulting network, the approach detects this and is protected
against over-fitting.

Future work on this topic could benefit from extending the FFBM to be hierarchical in nature. Such a modification would be relatively natural and would increase the descriptive power of the model. The current FFBM formulation is limited in that it can only explain structure at the macro-scale. Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.