### 2021-11-07 publication/abstract.tex
Flesh out inference

\begin{abstract}
Labelled networks are a very common and important class of data.
A typical inference goal is to determine how the vertex labels
(or {\em features}) affect the network's structure.

In this work, we introduce a new generative model, the feature-first block model (FFBM),
to facilitate the use of richer queries on labelled network structure.
We develop a Bayesian framework and devise a two-level Markov-Chain Monte-Carlo approach to efficiently sample the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.

We apply the proposed methods to a variety of network data
to extract the most important features along which the vertices
are partitioned. The main advantages of the proposed approach are
that the whole feature-space is used automatically and that features
can be rank-ordered implicitly according to impact.

\keywords{Labelled Networks \and Bayesian Inference.}

\end{abstract}