### 2021-11-16 publication/prelim.tex

\section{Preliminaries}

We first need a model for community-like structure in a network. For this we adopt the widely-used stochastic block model (SBM). This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical}. To allow for degree-variability between members of the same block, we adopt the following degree-corrected 
formulation (DC-SBM):
% , defined in~(\ref{defn:microcan-dc-sbm}).

\begin{definition}[Microcanonical DC-SBM]
	\label{defn:microcan-dc-sbm}
	Let $N \geq 1$ denote the number of vertices in an undirected graph
with $E$ edges. The block memberships are encoded by a vector $b \in [B]^N$,
where $B$ is the number of non-empty blocks.\footnote{For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.}
	Let $e=(e_{rs})$ be the $B \times B$ symmetric matrix of edge counts 
between blocks, such that $e_{rs}$ is the number of edges from block $r$ to 
block $s$. 
	Let $k =(k_i)$ denote a vector of length $N$, with $k_i$ being the degree of vertex $i$.

The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated 
by placing edges uniformly at random, conditional
on the constraints imposed by $b$, $e$ and $k$ being satisfied.
Specifically, if $A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k)$,
then with probability~1 it satisfies,
for all $r,s\in[B]$
and all $i\in[N]$:
	%
	\begin{equation}
		e_{rs} = \sum_{i, j \in [N]} A_{ij} 
	\one \{b_i = r\} \one \{b_j = s\} 
		\qquad 
		\textrm{and} \qquad
		k_i = \sum_{j \in [N]} A_{ij}.
		\label{eqn:sbm-constraints}
	\end{equation}
\end{definition}