### 2022-02-20 iwsm/references.tex

% References should be placed in the text in author (year) form.
% The list of references should be placed below IN ALPHABETICAL ORDER.
% (Please follow the format of the examples very tightly).

\references
\begin{description}

	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.} pp. 1–-8.

	\item[Airoldi, E.M., Blei, D., Fienberg, S., Xing, E.] (2009).
	Mixed membership stochastic blockmodels. 
	In: {\it Advances in Neural Information Processing Systems} vol. 21.

	\item[Andrieu, C., Roberts, G.O.] (2009). 
	The pseudo-marginal approach for efficient Monte Carlo computations.
	The Annals of Statistics 37(2), 697–-725.

	\item[Gaucher, S., Klopp, O., Robin, G.] (2021).
	Outliers detection in networks with missing links.
	Computational Statistics \& Data Analysis 164, 107308.

	\item[Hastings, W.K.] (1970).
	Monte Carlo sampling methods using Markov chains and their applications.
	Biometrika 57(1), 97-–109.

	\item[Leskovec, J., Mcauley, J.] (2012).
	Learning to discover social circles in ego networks.
	In: {\it Advances in Neural Information Processing Systems} vol. 25

	\item[Nowicki, K., Snijders, T.A.B.] (2001).
	Estimation and prediction for stochastic blockstructures. 
	Journal of the American Statistical Association 96(455), 1077–-1087.

	\item[Pasternak, B., Ivask, I.] (1970).
	Four unpublished letters. Books Abroad 44(2), 196–-200.

	\item[Peixoto, T.P.] (2014).
	Efficient Monte Carlo and greedy heuristic for the inference of stochastic block models.
	Physical Review E 89(1).

	\item[Peixoto, T.P.] (2017).
	Nonparametric Bayesian inference of the microcanonical
	stochastic block model. Physical Review E 95(1).

	\item[Peixoto, T.] (2014).
	The graph-tool Python library. figshare (2014), figshare. com/articles/graph\_tool/1164194

	\item[Roberts, G.O., Tweedie, R.L.] (1996).
	Exponential convergence of Langevin distributions and their discrete approximations.
	Bernoulli 2(4), 341–-363.

	\item[Stehle, J. et al] (2011).
	High resolution measurements of face-to-face contact patterns in a primary school.
	PLoS ONE 6(8), 1–-13.

	\item[Zhu, J., Song, J., Chen, B.] (2016).
	Max-margin nonparametric latent feature models for link prediction.
	arXiv preprint cs.LG:1602.07428.
\end{description}