### 2022-03-05 iwsm/references.tex

% References should be placed in the text in author (year) form.
% The list of references should be placed below IN ALPHABETICAL ORDER.
% (Please follow the format of the examples very tightly).

\references
\begin{description}

	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.}.

%	\item[Airoldi, E.M., Blei, D., Fienberg, S., Xing, E.] (2009).
%	Mixed membership stochastic blockmodels. 
%	In: {\it Advances in Neural Information Processing Systems} vol. 21.

	\item[Gaucher, S., Klopp, O., Robin, G.] (2021).
	Outliers detection in networks with missing links.
	Computational Statistics \& Data Analysis 164.

	\item[Leskovec, J., Mcauley, J.] (2012).
	Learning to discover social circles in ego networks.
	In: {\it Advances in Neural Information Processing Systems} vol. 25

	\item[Nowicki, K., Snijders, T.A.B.] (2001).
	Estimation and prediction for stochastic blockstructures. 
	Journal of the American Statistical Association 96(455), 1077–-1087.

	\item[Pasternak, B., Ivask, I.] (1970).
	Four unpublished letters. Books Abroad 44(2), 196–-200.

%	\item[Peixoto, T.P.] (2013).
%	The graph-tool Python library. figshare (2014), figshare. com/articles/graph\_tool/1164194

%	\item[Peixoto, T.P.] (2014).
%	Efficient Monte Carlo and greedy heuristic for the inference of stochastic block models.
%	Physical Review E 89(1).

	\item[Peixoto, T.P.] (2017).
	Nonparametric Bayesian inference of the microcanonical
	stochastic block model. Physical Review E 95(1).

%	\item[Roberts, G.O., Tweedie, R.L.] (1996).
%	Exponential convergence of Langevin distributions and their discrete approximations.
%	Bernoulli 2(4), 341–-363.

	\item[Stehle, J. et al] (2011).
	High resolution measurements of face-to-face contact patterns in a primary school.
	PLoS ONE 6(8), 1–-13.

%	\item[Zhu, J., Song, J., Chen, B.] (2016).
%	Max-margin nonparametric latent feature models for link prediction.
%	arXiv preprint cs.LG:1602.07428.
\end{description}