### 2022-03-20 iwsm/abstract.tex
Integrate review comments

\abstract{
Labelled networks are an important class of data,
naturally appearing
in applications in science and engineering.
A typical inference goal is to determine how the vertex labels
(or {\em features}) affect the network's structure.
We introduce a new generative model, the feature-first block model (FFBM),
that facilitates the use of rich queries on labelled networks.
We develop a Bayesian framework and devise a two-level Markov chain Monte 
Carlo approach to efficiently sample from the
posterior distribution of the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.
We apply the proposed methods to several real-world networks
to extract the most important features along which the vertices
are partitioned. Importantly, the whole feature-space is used automatically and features
can be rank-ordered implicitly by importance.
[The full version of this paper is available on arXiv as [cs.LG] 2105.13762.]
}