### 2022-05-22 abstract_IWSM2022_Tray_final/references.tex

% References should be placed in the text in author (year) form.
% The list of references should be placed below IN ALPHABETICAL ORDER.
% (Please follow the format of the examples very tightly).

\references
\begin{description}

	\item[Krebs, V.] (2004) . The political books network,
	http://www.orgnet.com/.

	\item[Leskovec, J., Mcauley, J.] (2012).
	Learning to discover social circles in ego networks.
	{\it Advances in Neural Information Processing Systems} vol. 25

	\item[Peixoto, T.P.] (2017).
	Nonparametric Bayesian inference of the microcanonical
	stochastic block model. Physical Review E 95(1).

	\item[Roberts, G.O., Tweedie, R.L.] (1996).
	Exponential convergence of Langevin distributions and their discrete approximations.
	Bernoulli 2(4).

	\item[Stanley, N. et al] (2019). Stochastic block models with multiple continuous attributes. Applied Network Science, 4(1), 1-22.

	\item[Stehle, J. et al] (2011).
	High resolution measurements of face-to-face contact patterns in a primary school.
	PLoS ONE 6(8), 1–-13.

	\item[Tray, L., Kontoyiannis, I.] (2021).
	The feature-first block model.
	arXiv preprint 2105.13762 [cs.LG].

\end{description}